Increased Goal Scoring in the NHL: An Attempt To Expand Popularity

### 2023-04-19 outputs/paper/paper.qmd
new input data, changed scripts, paper updates

---
title: "Expanding the Popularity of the NHL Through Increased Goal Scoring"
subtitle: "An Analysis of the Increase in Goal Scoring From 2011 to 2023"
author: 
  - Jason Ngo
thanks: "Code and data are available at: https://github.com/thejasonminh/nhlpoints"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "Of the four major sports in the USA, the NHL is the least popular. In an effort to grow the game and make the game more exciting, the NHL wanted to increase goal scoring. Using data from the NHL, the paper will use graphs and tables to analyze average leaguewide goal scoring per game from the 2011-2012 to 2022-2023 seaons. A linear regression model will also be created to predict goal scoring averages in the upcoming seasons. Results show that the NHL was succesful in their mission as the average goals per game increased per season, with the current season exhibiting the highest average goals per game in a 10 year period. These findings suggest that factors such as increased player talent, more penalties and rule changes have resulted in more goals scored every season. To further grow the game, analysis on the topic of matchups, streaming, and media narratives is recommended."
number-sections: true
bibliography: references.bib
---

@@

@@

# Data {#sec-data}
[@citeR]

@@

@@

>>>>

---
title: "Increased Goal Scoring in the NHL: An Attempt To Expand Popularity"
subtitle: "An Analysis of the Increase in Goal Scoring From 2011 to 2023"
author: 
  - Jason Ngo
thanks: "Code and data are available at: https://github.com/thejasonminh/nhlpoints"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "Of the four major sports in the USA, hockey is the least popular. In an effort to grow popularity, the NHL (National Hockey League) strove to increase goal scoring. Using data from the NHL, the paper will analyze average leaguewide goal scoring per game from the 2011-2012 to 2022-2023 seaons. A linear regression model will also be created to predict goal scoring averages in the upcoming seasons. Results show that the NHL was succesful as the average goals per game increased over a 10 year period. These findings suggest that increased player talent, more penalties and rule changes have resulted in more goals."
number-sections: true
bibliography: references.bib
---

@@

rework this in the context of goal scoring

@@

# Data {#sec-data}
talk about league dataset, forwards and dmen dataset
talk about merging the nhl shooting pct into the league dataset
talk about why i didnt choose to use nhl.com for league dataset(annoying to work with 12 datasets)
talk about why i chose the columns i chose
talk about why i chose the ppg values for forwards and dmen
relationship between save pct, gaa, goals/game and shooting pct

data is collected real time every game, with internal models used to calculate the special stats
[@citeR]

@@

12 season trends in league data
increase in good forwards and defensemen

@@

goal scoring did increase, why?
player talent increased -> correlates to better powerplays
rule changes = less powerplays, could effect even strength scoring
goalie equipment changing

weaknesses = i did not go into an extreme in-depth level of the players themselves. there is a whole world of advance stats that could be used to analyze the increase of player talent.

next steps = analyze viewership data to see the effect goal scoring has had on it. if we want to focus more on growing game, analysis on rivalries, game streaming and media narratives should be done.
